---
title: "Study on US Weather Events that Impact Health and Economy"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---


## Data Processing

The raw data can be found on github or https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2

The code below shows the data processing.

```{r cache = TRUE}

storm_df <- read.csv("repdata_data_StormData.csv.bz2")
```

Getting the data for population health. I combined the fatalities and injuries by 
multiplying the fatality by 10 and then adding it to the number of injuries, and then
averaging it over each weather event.
```{r cache = TRUE}

library(dplyr)
health_df <- select(storm_df, EVTYPE, FATALITIES, INJURIES)
health_df <- health_df %>% mutate(COMBINED=10*FATALITIES + INJURIES)
health_res <- aggregate(health_df$COMBINED,by=list(Event=health_df$EVTYPE), FUN="mean")

```


Getting the data for economic impact.
```{r cache = TRUE}
econ_df <- select(storm_df, EVTYPE, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP)
# Scale based on the PROPDMGEXP
econ_df$PROPDMGEXP[econ_df$PROPDMGEXP == "K"] <- 1000
econ_df$PROPDMGEXP[econ_df$PROPDMGEXP == "M"] <- 1000000
econ_df$PROPDMGEXP[econ_df$PROPDMGEXP == "B"] <- 1000000000
econ_df$PROPDMGEXP[econ_df$PROPDMGEXP == ""] <- 0
econ_df$CROPDMGEXP[econ_df$CROPDMGEXP == "K"] <- 1000
econ_df$CROPDMGEXP[econ_df$CROPDMGEXP == "M"] <- 1000000
econ_df$CROPDMGEXP[econ_df$CROPDMGEXP == "B"] <- 1000000000
econ_df$CROPDMGEXP[econ_df$CROPDMGEXP == ""] <- 0
  
econ_df$PROPDMGEXP <- as.numeric(econ_df$PROPDMGEXP)
econ_df$CROPDMGEXP <- as.numeric(econ_df$CROPDMGEXP)

econ_df <- econ_df %>% mutate(COST= PROPDMG*PROPDMGEXP + CROPDMG*CROPDMGEXP)
econ_res <- aggregate(econ_df$COST,by=list(Event=econ_df$EVTYPE), FUN="mean")

```



## Results


```{r}

boxplot(health_res$x, main="Box plot of combined fatality and injuries", 
        ylab="Fatality+Injury Impact")
```

The weather event that is most harmful to the population health is: 
```{r} 
print(health_res[which.max(health_res$x),1])
```


```{r}

boxplot(econ_res$x, main="Box plot of economic cost", 
        ylab="Property and Crop Damage")

```

The weather event that has the greatest economic consequence is: 
```{r} 
print(health_res[which.max(econ_res$x),1])
```

